Visualizing Project Evolution Through Abstract Syntax Tree Analysis

### 2016-07-11 paper/document.tex
Cut down to 10 pages.

Visualizations can assist our understanding of a software repository. The history of a code base is often tracked in a version control system such as Git or SVN. Popular source code repository hosting websites such as GitHub~\cite{github-graphs} and BitBucket~\cite{bitbucket-graphs} often provide a way to visualize project evolution. These visualizations often show changes to lines of code, as line changes are what the underlying version control system tracks. The advantage of visualizing line changes is that it works for any text-based file. There is a major disadvantage, however: by only visualizing changes to lines of code, GitHub (Figure~\ref{fig:GH_code_freq}) and similar visualizations lose the \emph{meaning} behind those changes. Although a programmer may type individual characters into their text editor, the lines of code represent something much deeper. A Java programmer, for instance, writes lines of code with the \emph{intent} of creating packages, defining classes, declaring fields, and implementing methods. An author commits lines of code to implement features, write tests, fix bugs, and refactor code bases. Lines of code merely tell us that changes were made; not the \emph{reason} for the changes.

@@

\begin{figure}[tb]
\centering
\includegraphics[width=\columnwidth]{data-extracted}
\caption{Hierarchy of data extracted by TypeV. Each repository may have zero or more commits; each commit may have zero or more changed files; each changed files may have zero or more changes, as computed by our AST differencing tool.}
\label{fig:data-extracted}
\end{figure}

@@

In Java, a type can be primitive, provided by the language itself, or an object reference. Primitive types are those that are built into the language such as an \texttt{int} or \texttt{char}. Object references point to instances of classes. Java's syntax encourages the creation of classes, thus we expect substantial Java projects to define many domain-specific types. Additionally, Java places the constraint that each source file must define one class. When writing or contributing to a project, the author's choice of which part of the program to edit will in turn affect what types they end up working with --- that is, it will affect the author's type coverage. Similarly, the set of files that an author works with affects that author's file coverage.

@@

Internal threats to validity include our treatment of merge conflicts. As mentioned before, we ignore all merges, so if there is a merge conflict, it is possible that a developer makes non-trivial changes. Changes made during merge conflict resolution are missed by TypeV. We have made the assumptions that if this case happens, the changes are minor.

@@

A weakness of the timeline view described in this paper is that details about an author's activities are buried. As future work, the data extracted by TypeV can be used to create a visualization similar to Gource~\cite{gource} or code\_swarm~\cite{codeswarm}. These visualizations animate authors scurrying back-and-forth over files they edit; we proposed to animate authors' activities over types. The focal point would be a particular type or group of types. The type itself may be represented abstractly as a shape or its UML representation that grows and shrinks as its usage increases and wanes over time.

@@

\section*{Acknowledgments}

The authors are grateful to the reviewers for their helpful suggestions on improving this work.

>>>>

Visualizations can assist our understanding of a software repository. Popular source code repository hosting websites such as GitHub~\cite{github-graphs} and BitBucket~\cite{bitbucket-graphs} often provide a way to visualize project evolution. These visualizations often show changes to lines of code, as line changes are what the underlying version control system tracks. The advantage of visualizing line changes is that it works for any text-based file. There is a major disadvantage, however: by only visualizing changes to lines of code, GitHub (Figure~\ref{fig:GH_code_freq}) and similar visualizations lose the \emph{meaning} behind those changes. Although a programmer may type individual characters into their text editor, the lines of code represent something much deeper. A Java programmer, for instance, writes lines of code with the \emph{intent} of creating packages, defining classes, declaring fields, and implementing methods. An author commits lines of code to implement features, write tests, fix bugs, and refactor code bases. Lines of code merely tell us that changes were made; not the \emph{reason} for the changes.

@@

\begin{figure}[tb]
\centering
\includegraphics[width=\columnwidth]{data-extracted}
\caption{Hierarchy of data extracted by TypeV. Each repository may have zero or more commits; each commit may have zero or more changed files; each changed file may have zero or more changes, as computed by our AST differencing tool.}
\label{fig:data-extracted}
\end{figure}

@@

Java's syntax encourages the creation of classes, thus we expect substantial Java projects to define many domain-specific types. Additionally, Java places the constraint that each source file must define one class. When writing or contributing to a project, the author's choice of which part of the program to edit will in turn affect what types they end up working with --- that is, it will affect the author's type coverage. Similarly, the set of files that an author works with affects that author's file coverage.

@@

Internal threats to validity include our treatment of merge conflicts. As mentioned in section~\ref{sec:methodology}, we ignore all merges, so if there is a merge conflict, it is possible that a developer made non-trivial changes. Changes made during merge conflict resolution are missed by TypeV. We have made the assumptions that if this case happens, the changes are minor.

@@

A weakness of the timeline view described in this paper is that details about an author's activities are buried. As future work, the data extracted by TypeV can be used to create a visualization similar to Gource~\cite{gource} or code\_swarm~\cite{codeswarm}. These visualizations animate authors moving back-and-forth over files they edit; we proposed to animate authors' activities over \emph{types}. The focal point would be a particular type. The type itself may be represented abstractly as a shape or its UML representation that grows and shrinks as its usage increases and wanes over time.

@@

%\section*{Acknowledgments}

%The authors are grateful to the reviewers for their helpful suggestions on improving this work.
